---
title: "Startup Funds Investigation"
date: "July 15, 2015"
output: 
  knitrBootstrap:::simple_document:
    theme: cerulean
    highlight: idea
    menu: FALSE
    clean_supporting: TRUE
    toc: TRUE
---


```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(htmltools)
library(stringr) # for string manipulation
library(lubridate) # for date manipulation
library(reshape2) # for reshaping data
library(ggplot2)
library(scales)
library(dplyr) # for %>% operator
library(DT)
# s0 <- knitr::knit_hooks$get('source')
# knitr::knit_hooks$set(
#   list(
#     source=function(x,options){
#       if(is.null(options$class)) s0(x, options)
#       else
#         tags$button(class=c())
#     }
#   )
# )
# 
knitr::opts_chunk$set(bootstrap.show.code=FALSE, bootstrap.show.output=FALSE, bootstrap.show.message=FALSE, warning=F, message=F, out.width="100%", cache=T, echo=F, dpi=300)
knitr::opts_knit$set(root.dir="../../IowaStateStartupFunds/")

# Transformation - log(x+1) to deal with zero data 
log_one_trans <- function() trans_new("log_one", function(x) log10(x+1), function(x) 10^x-1, domain=c(0, Inf))
```

# Introduction

The following information was taken from meeting notes (meetings on 6/4/2015 and 5/20/2015)

## Goals for the project

1. Examine trends (or lack thereof) in startup funding over time
2. Examine return on investment for startup funding 
    - Publications
    - Proposals submitted
    - Grants received
    - Tenure?

## Data
```{r data-description, echo=F, results='asis', cache=F}
library(htmltools)
HTML(paste0(readLines("Writing/DataDescriptions.html")))
```

## Other factors

- 2003-2008 was a competitive hiring season (may have larger startup packages as a result)
- 2008 Crash and budget freeze at the state level
- American Recovery and Reinvestment Act (2009) with increased funding levels
- 2012 slowdown in research funding (may need to compare to NSF/NIH/USDA/DOD/DOE/Ames Lab funding levels)


# Analysis of Individual Datasets

```{r OVPRdata, cache=F}
# source("Code/CleanStartupFundingData.R")
load("Data/Hires.RData")
load("Data/FundingStartup.RData")
```

```{r OSPAdata, cache=F}
# source("Code/CleanAwardsData.R")
# source("Code/CleanProposalsData.R")
load("Data/Awards.RData")
load("Data/Proposals.RData")
```

```{r MergeDatasetsSetup, cache=F}
received.money <- awards %>% 
  group_by(Investigator) %>% 
  summarize(grant.received = sum(Award.Amount), 
            n.awards = length(unique(Award.Number)))
received.money$grant.received[is.na(received.money$grant.received)] <- 0

submitted.proposals <- prop %>% 
  group_by(Investigator) %>% 
  summarize(total.proposal.money = sum(Proposal.Amount), 
            n.proposals = length(unique(Proposal.Number)), 
            prop.fund.pct = mean(Proposal.Status=="Funded"))
submitted.proposals$total.proposal.money[is.na(submitted.proposals$total.proposal.money)] <- 0

isuHires <- merge(hires, received.money, by.x="Name", by.y="Investigator", all.x=T, all.y=F)
isuHires <- merge(isuHires, submitted.proposals, by.x="Name", by.y="Investigator", all.x=T, all.y=F)
isuHires$grant.received[is.na(isuHires$grant.received)] <- 0
isuHires$total.proposal.money[is.na(isuHires$total.proposal.money)] <- 0

isuHires$ProfRankFac <- isuHires$ProfRank %>% 
  str_replace("(Adjunct/Lecturer)|(Affiliate Prof)|(Clinician)", "Other") %>%
  str_replace("Distg ", "") %>%
  factor(levels=c("Asst Prof", "Assoc Prof", "Prof", "Other"), labels=c("Assistant", "Associate", "Professor", "Other"))

isuHires$CollegeFac <- isuHires$College %>% factor(labels=c("AGLS", "Bus", "Des", "Engr", "Hum Sci", "LAS", "Vet Med"))
isuHires$CollegeFac <- as.character(isuHires$CollegeFac)
isuHires$CollegeFac[isuHires$CollegeFac=="LAS" & isuHires$Dept1%in%c("Anthro", "Econ", "Greenlee", "Pol Sci", "Psych", "Soc")] <- "LAS: Social Sci"
isuHires$CollegeFac[isuHires$CollegeFac=="LAS" & isuHires$Dept1%in%c("BBMB", "Chem", "Com S", "EEOB", "GDCB", "GeAt", "Math", "Physics", "Stat")] <- "LAS: Science"
isuHires$CollegeFac[isuHires$CollegeFac=="LAS" & isuHires$Dept1%in%c("Engl", "FL&L", "Hist", "Music", "Phil & RS", "WL&C")] <- "LAS: Arts"

isuHires$YearsAtISU <- 2015 - isuHires$Year

isuHires$YearsAtISUFac <- cut_number(isuHires$YearsAtISU, n=4)

# Quartiles for startup funding, by college
startup.quantiles <- isuHires %>% 
  group_by(CollegeFac) %>% 
  summarize(Q1=quantile(Total.Cost, .25, type=8), 
            Q2=quantile(Total.Cost, .5, type=8), 
            Q3=quantile(Total.Cost, .75, type=8), 
            Q4=quantile(Total.Cost, 1, type=8)) %>% 
  melt(id.vars=1, measure.vars=2:ncol(.), variable.name="Quantile", value.name="Total.Cost")

# Rank individual faculty startup packages by the quartile of their respective college. 
isuHires <- isuHires %>% 
  group_by(CollegeFac) %>%
  mutate(TotalCostFac = as.numeric(cut(Total.Cost, breaks=c(-1, quantile(Total.Cost, probs=1:4/4)))), 
         TotalCostDecile = as.numeric(cut(Total.Cost, breaks=c(-1, quantile(Total.Cost, probs=1:10/10)))))
isuHires$TotalCostFac[is.na(isuHires$TotalCostFac)] <- 1
isuHires$TotalCostDecile[is.na(isuHires$TotalCostDecile)] <- 1
isuHires$TotalCostFac <- factor(isuHires$TotalCostFac, levels=1:4, labels=c("Startup $: Q1", "Startup $: Q2", "Startup $: Q3", "Startup $: Q4"))
isuHires$TotalCostDecile <- factor(isuHires$TotalCostDecile, levels=1:10)
```

## OVPR Data
A full exploration of the data from the Office of the Vice President for Research can be found [here](OVPR.html)

## OSPA Data
A full exploration of the data from the Office of Sponsored Programs Administration can be found [here](OSPA.html)

## Google Scholar Data
Results from Google Scholar were aggregated between July 6 and July 12, 2015. Google search results (and corresponding BibTeX citations) were collected for results as follows:

Query | Criteria/Example
----- | --------
Author Name | "Smith, John"
Date | Year>=2005
Text Contains | "Iowa State University"

Results were also filtered to ensure that the results were in areas that John Smith at Iowa State would publish in (where possible, ambiguities were resolved by referring to an online CV). 

```{r GScholarData, cache=F}
# source("Code/Publication Results.R")
load("Data/Publications.RData")
```
```{r GScholarMerge, cache=F}
# Hire information
tmp <- isuHires[,c("Name", "CollegeFac", "Dept1", "Dept2", "Year", "ProfRank", "ProfRankFac", "Total.Cost", "TotalCostFac")]
tmp$Name <- str_to_title(tmp$Name) %>% str_replace(fixed("O'd"), "O'D")
citations <- merge(tmp, citations, by.x="Name", by.y="author", all=T)
names(citations)[names(citations)=="year"] <- "year.pub"

# Add in extra years
citations <- citations %>% ungroup %>% group_by(Name) %>% do({
  idx <- !c(2005:2015)%in%.$year.pub
  if(sum(idx)>0){
    newdf <- data.frame(unique(.[,c("Name", "CollegeFac", "Dept1", "Dept2", "Year", "ProfRank", "ProfRankFac", "Total.Cost", "TotalCostFac")]), author.match=unique(.$Name), year.pub=c(2005:2015)[idx])
    rbind_list(., newdf) %>% as.data.frame() %>% arrange(year.pub)
  } else {
    .
  }
}) 

citations$YearsAtISU <- with(citations, year.pub-Year)

# Remove citations from more than one year prior to hire
# citations <- subset(citations, year.pub+1>=Year)

citations.yearly <- citations %>% 
  group_by(Name, CollegeFac, Dept1, Year, year.pub, ProfRank, ProfRankFac, Total.Cost, TotalCostFac) %>% 
  summarize(
    YearsAtISU=unique(year.pub-Year),
    total.pubs=length(type[!is.na(type)]), 
    articles=sum(type=="article", na.rm=T), 
    books=sum(type=="book", na.rm=T), 
    other=sum(type=="other", na.rm=T))
citations.yearly$atISU <- with(citations.yearly, year.pub-Year>=0)


citations.total <- citations %>% 
  group_by(Name, CollegeFac, Dept1, Year, ProfRank, ProfRankFac, Total.Cost, TotalCostFac) %>%   
  summarize(
    YearsAtISU=2015-unique(Year),
    total.pubs=length(type[!is.na(type)]), 
    articles=sum(type=="article", na.rm=T), 
    books=sum(type=="book", na.rm=T), 
    other=sum(type=="other", na.rm=T))

rm(tmp)
```

```{r GScholarCollege, fig.width=6, fig.height=4, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.pubs), data=subset(citations.yearly, atISU & !is.na(CollegeFac))) +
  geom_vline(xintercept=0) + 
  geom_jitter(size=1, shape=1, position=position_jitter(width=.4, height=.4), alpha=.3) + 
  geom_smooth(se=F, span=.6, size=2, method="loess") + 
  facet_wrap(~CollegeFac, ncol=4, scales="free_y") + 
  xlab("Years at ISU") + ylab("Number of Yearly Publications") + 
  ggtitle("Yearly Publications") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(-5, 0, 5, 10))
```

Above, the x-axis shows time at Iowa State (or rather, time since hired), and the y-axis shows the yearly publications (books, articles, etc.). The blue lines indicate college averages over time (loess smooth). From this broad overview, we see that publications in many colleges peak at approximately 5 years, and then tail off slightly by year 10. 

Further analysis is provided in the Merged Data section. 

# Analysis of Merged Data

## Awarded Grant Funding
We'll begin by looking at the timeline - how many years at ISU are necessary before grant applications are successful? We split this by college and faculty rank (as we'd expect that full professors who are hired should be able to command grant money sooner than new faculty). Subsequent graphs show data from hired professors (i.e. not adjuncts, affiliates, lecturers, or clinicians). 

```{r YearsAtISUGrants, fig.width=6, fig.height=8, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received), data=subset(isuHires, ProfRankFac!="Other" )) +
  geom_jitter(shape=1, size=4, alpha=.75) + 
#   # geom_smooth(se=F, adjust=2) + 
#   scale_shape_manual(values=c('A', 'B', 'D', 'E', 'H', 'L', 'V'), guide=F) +
#   scale_color_discrete(guide=F) + 
  # guides(shape=guide_legend(ncol=4), color=guide_legend(ncol=4)) + 
  facet_grid(CollegeFac~ProfRankFac, scales="free") + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 1e3, 1e5, 1e7))
```
The most noticeable difference between colleges is that in engineering, assistant professors get grants about five years after they are hired, and it is very rare for professors at that stage in their careers not to get grant funding. This trend is present (though less pronounced) in Ag & Life Sciences and Human Sciences as well, though some assistant professors in those colleges seem to get funding much earlier in their careers (perhaps because of large grants given to groups of professors). In Liberal Arts and Sciences, grant funding does seem to increase with experience for some hires, but perhaps because LAS includes both arts (which are less grant-reliant) and sciences, this trend is less pronounced than in other colleges. 

In order to examine the effect of startup funding on grant receipts over a career of variable duration, we will examine the distribution of startup funding by college, categorizing startup costs by the quartile (calculated for the hire's college). Below, vertical lines mark these quantiles

```{r TotalCostQuartiles, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=Total.Cost, fill=CollegeFac), data=isuHires) + 
  geom_density(aes(y=..density..), origin=0, color="black", trim=T) +
  geom_vline(aes(xintercept=Total.Cost), data=startup.quantiles) + 
  scale_fill_discrete(guide=F) + 
  scale_x_continuous(trans='sqrt', breaks=c(1000, 15000, 50000, 100000, 250000, 500000, 1e6, 2e6), labels=c("1K", "15K", "50K", "100K", "250K", "500K", "1M", "2M"), limits=c(0, 2e6)) + 
  facet_grid(CollegeFac~.) + 
  ggtitle("Distribution of Startup Costs by College") +
  xlab("Startup Package Cost") + ylab("Count") + 
  # guides(fill=guide_legend(ncol=4)) + 
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical")
```

Using these quantile calculations, we then can examine total grants received as explained by years since hired. 

```{r YearsTotalCostGrantsByCollege, fig.width=8, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  geom_smooth(method="lm", span=.5, se=F, fullrange=F, aes(x=YearsAtISU, y=grant.received)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~CollegeFac, ncol=5) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) +
  scale_x_continuous(breaks=c(0, 5, 10))
```

It is important to note several caveats at this point: first, hires are not separated out by faculty rank or position, so hires whose primary responsibilities include administrative work may command large startup packages but bring in relatively little grant funding. Additionally, different colleges have different funding structures: LAS, for instance, may need to provide lab startup funding for science based departments, but would not have to provide this funding for new hires in English or History. Finally, the x-axis shows years since hire, which does not necessarily translate to years at ISU (faculty hired in 2005 may have left in 2011 after receiving tenure, for example). 

Linear regression lines are provided here as (extremely) rough approximations; it may be useful to examine the right-most endpoint (rather than the entire line) as a prediction of total grant money received after 10 years at ISU. 

With these caveats in mind, 

* Startup funding in LAS is correlated with later grant funding; this is likely explained by departmental startup funding (and grant funding) norms. 
* Startup funding in engineering is largely not predictive of grant receipts; after 10 years, hires with small startup packages receive approximately the same grant funding as hires with large startup packages. 
* Startup funding in Human Sciences is correlated with later grant funding- this may again be a departmental norm situation (food scientists may require lab space and compete for NSF/NIH grants, where professors in the hospitality management department require fewer financial resources when hired and receive fewer grants.)
* Grants are relatively rare in the business college.   

### Individual Colleges: Grant Proposals and Startup Funding by Department
Examining the LAS results in more detail, we see that departmental startup packages largely fall in the same quantiles. In Chemistry, Physics, BBMB, EEOB, Computer Science, Statistics, Economics, and Math, the career trajectory of grant proposals suggests that in most cases large startup packages are worthwhile investments. Psychology is more of a mixed bag; many psychologists received large startup packages but have not received significant grant funding. Disciplines with more economical startup packages are also less likely to have faculty receiving large grants; this is likely due to discipline norms, rather than faculty quality. 
```{r YearsTotalCostGrantsLAS-Arts, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="LAS: Arts" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date\n(LAS: Arts)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```
```{r YearsTotalCostGrantsLAS-Science, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="LAS: Science" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date\n(LAS: Science)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```
```{r YearsTotalCostGrantsLAS-SocialScience, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="LAS: Social Sci" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date\n(LAS: Social Science)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```
By comparison, the College of Engineering is fairly consistent; some departments have more startup funding (MSE, CBE), but overall, grant receipts are largely similar (though there is not sufficient data for Ag & Biosystems Engineering to determine whether that department receives grants at the same rate as other departments). Grant funding may increase slightly over the course of a career (there are increases in AerE, MSE, and ME that are fairly consistent). After 2 years, most hires will have received at least $100,000 in grant funding, and many hires will have received at least one million in cumulative grants after 5-6 years. Startup funding does not seem to affect these trends in a significant way, except that more recent hires seem more likely to receive large startup packages, and are less likely to have received grant money at this point in their careers.  
```{r YearsTotalCostGrantEngr, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Engr" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

Examining the College of Agriculture & Life Sciences, we find again that different departments have different startup funding resources, but that within a department, startup funding does not seem to greatly predict later grant receipts. 
```{r YearsTotalCostGrantAGLS, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="AGLS" & ProfRankFac!="Other" & Dept1!="Stat")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

Examining the College of Human Sciences, we find again that different departments have different startup funding resources; Food Science and Human Nutrition in particular tends to have more startup funding than other departments, while the School of Education has a uniquely time-dependent startup funding allocation - hires from FY06-FY10 are typically in the first quartile of startup package funding, while hires from FY11-FY13 are typically in the second quadrant, and hires from FY13-FY15 are typically in the third quadrant. 
```{r YearsTotalCostGrantHumSci, fig.width=6, fig.height=6, dpi=300}
isuHires$YearsAtISUjit <- jitter(isuHires$YearsAtISU)
ggplot(aes(x=YearsAtISUjit, y=grant.received), data=subset(isuHires, CollegeFac=="Hum Sci" & ProfRankFac!="Other")) +
  geom_point(aes(shape=TotalCostFac, color=TotalCostFac), size=3) +
  # geom_point(aes(color=TotalCostFac), data=subset(isuHires, CollegeFac=="Hum Sci" & !ProfRankFac%in%c("Assistant", "Other")), size=5, shape=1, show_guide=F) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
isuHires <- isuHires[,-which(names(isuHires)=="YearsAtISUjit")]
```

Examining the College of Veterinary Medicine, we find great differences in grant receipts between departments; grants are almost unheard of in Veterinary Pathology, but are reasonably common in Veterinary Diagnostic and Production Animal Medicine. In this college, startup funding does seem to somewhat correlate with grant receipts; the most expensive startup packages typically receive grant funding, particularly in the BMS, VCS, and VDPAM departments. 
```{r YearsTotalCostGrantVetMed, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Vet Med" & ProfRankFac!="Other")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

Examining the College of Business, we find that grant funding is so rare in business that it would be nearly impossible to draw meaningful conclusions about the relationship between startup funding and grants. Additionally complicating the issue, funding for startup packages seems to be higher for hires in FY11-FY16 than in FY06-FY10. 
```{r YearsTotalCostGrantBus, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Bus" & ProfRankFac!="Other")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```


Examining the College of Design, we find again that startup costs do not seem to significantly influence later grant receipts once we account for department and time since hiring. 
```{r YearsTotalCostGrantDes, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Des" & ProfRankFac!="Other")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

## Proposed Grant Funding

```{r YearsAtISUProposals, fig.width=6, fig.height=8, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money), data=subset(isuHires, ProfRankFac!="Other" )) +
  geom_jitter(shape=1, size=4, alpha=.75) + 
  # geom_smooth(se=F, adjust=2) + 
  # guides(shape=guide_legend(ncol=7), color=guide_legend(ncol=7)) + 
  facet_grid(CollegeFac~ProfRankFac, scales="free") + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Proposed funding based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 1e3, 1e5, 1e7))
```
The most noticeable difference between colleges is that in Engineering and AGLS, assistant professors submit proposals immediately. Most, but not all, Human Sciences professors also begin submitting grants almost immediately after they are hired. In Liberal Arts and Sciences, grant funding does seem to increase with experience for some hires, but perhaps because LAS includes both arts (which are less grant-reliant) and sciences, this trend is less pronounced than in other colleges and some professors (at all ranks) do not ever apply for grant funding. In the business college, it is incredibly rare for professors to apply for grants, and it is common that professors do not apply for grants in the design college. 

As in the previous section, we will use quartiles to categorize startup funding by college so that we can examine total grants received as explained by years since hired. 

```{r YearsTotalCostProposalsByCollege, fig.width=8, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~CollegeFac, ncol=5) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

It is important to note several caveats at this point: first, hires are not separated out by faculty rank or position, so hires whose primary responsibilities include administrative work may command large startup packages but bring in relatively little grant funding. Additionally, different colleges have different funding structures: LAS, for instance, may need to provide lab startup funding for science based departments, but would not have to provide this funding for new hires in English or History. Finally, the x-axis shows years since hire, which does not necessarily translate to years at ISU (faculty hired in 2005 may have left in 2011 after receiving tenure, for example). 
With these caveats in mind, 

* Startup funding in LAS is highly correlated with later grant proposals; this is likely explained by departmental startup funding (and grant funding) norms. This correlation is much stronger than the correlation between startup funding and grant awards.
* Startup funding in Engineering is not very well correlated with proposed funding; this is largely because proposed funding is fairly consistent across engineering career trajectories. It may be the case that once a grant is successful, engineering professors do not devote much time to applying for additional funding. 
* In Human Sciences, more than in any other college, startup funding seems to be primarily dependent on the year of hire. Grant proposals are fairly common across the range of startup packages. 
* Grant proposals are relatively rare in the business college.   

### Individual Colleges: Proposals and Startup Funding by Department

Examining the LAS results in more detail, we see that departmental startup packages largely fall in the same quantiles. In Chemistry, Physics, BBMB, EEOB, Computer Science, Statistics, Psychology, and Math, the career trajectory of grant proposals suggests that the large startup packages are potentially worthwhile investments; these disciplines are also largely in the top two quartiles for startup funding. While many psychologists received large startup packages but have not received significant grant funding, it is apparent from this result that psychologists are applying for large amounts of funding, but are not necessarily successful. Disciplines with more economical startup packages are also less likely to have faculty applying for large grants; this is likely due to discipline norms, rather than faculty quality. 
```{r YearsTotalCostProposalsLAS-Arts, cache=T, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="LAS: Arts" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date (LAS: Arts)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

```{r YearsTotalCostProposalsLAS-Science, fig.width=6, fig.height=6, dpi=300, dependson='TotalCostQuartiles'}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="LAS: Science" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date\n(LAS: Science)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```
```{r YearsTotalCostProposalsLAS-SocialScience, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="LAS: Social Sci" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date\n(LAS: Social Science)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```
By comparison, the College of Engineering is fairly consistent; some departments have more startup funding (MSE, CBE), but overall, grant proposals are largely similar (though there is not sufficient data for Ag & Biosystems Engineering to determine whether that department submits proposals for grants of the same magnitude as other departments). Proposed funding does not seem to increase significantly over the course of a career, in contrast to the rate of successful proposals. After 2 years, most hires will have proposed at least $100,000 in grant funding, and many hires will have proposed at least one million in cumulative grants after 5-6 years. 
```{r YearsTotalCostProposalsEngr, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Engr" & ProfRankFac!="Other" )) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

Examining the College of Agriculture & Life Sciences, we find again that different departments have different startup funding resources, but that within a department, startup funding does not seem to greatly predict the amount of proposed funding. 
```{r YearsTotalCostProposalsAGLS, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="AGLS" & ProfRankFac!="Other" & Dept1!="Stat")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

Examining the College of Human Sciences, we notice that the previous disparity in grant funding within the school of education does not appear to be due to differences in funding proposals submitted; rather, faculty hired with tenure (circled values) seem to be more successful at proposing grants which are funded. 
```{r YearsTotalCostProposalsHumSci, fig.width=6, fig.height=6, dpi=300}
isuHires$YearsAtISUjit <- jitter(isuHires$YearsAtISU)
ggplot(aes(x=YearsAtISUjit, y=total.proposal.money), data=subset(isuHires, CollegeFac=="Hum Sci" & ProfRankFac!="Other")) +
  geom_point(aes(shape=TotalCostFac, color=TotalCostFac), size=3) +
  geom_point(aes(color=TotalCostFac), data=subset(isuHires, CollegeFac=="Hum Sci" & !ProfRankFac%in%c("Assistant", "Other")), size=5, shape=1, show_guide=F) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
isuHires <- isuHires[,-which(names(isuHires)=="YearsAtISUjit")]
```

Examining the College of Veterinary Medicine, we find that proposals may be somewehat correlated with startup funding in the VCS department, but that this correlation does not hold up in other departments. In addition, funding proposals in the VDPAM department seem to be more common for those hired between FY11-FY16 than for those hired in earlier years. 
```{r YearsTotalCostProposalsVetMed, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Vet Med" & ProfRankFac!="Other")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

Examining the College of Business, we find that grant proposals are so rare in business that it would be nearly impossible to draw meaningful conclusions about the relationship between startup funding and proposed grant funding. Additionally complicating the issue, funding for startup packages seems to be higher for hires in FY11-FY16 than in FY06-FY10. 
```{r YearsTotalCostProposalsBus, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Bus" & ProfRankFac!="Other")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```


Examining the College of Design, we find that startup costs do not seem to significantly influence funding proposals. 
```{r YearsTotalCostProposalsDes, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.proposal.money, shape=TotalCostFac, color=TotalCostFac), data=subset(isuHires, CollegeFac=="Des" & ProfRankFac!="Other")) +
  geom_jitter(size=3, position=position_jitter(width=.5, height=0)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Proposed $") + 
  ggtitle("Grant proposals based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7, 1e8), limits=c(0, 3e8)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

## Publications and Startup Costs

We initially consider the effect that startup packages have on the number of publications (as measured by results aggregated in Google Scholar). Publications here include journal articles, books, conference proceedings (such as IEEE) and miscellaneous other research articles. 


```{r GScholarCollegeStartup, fig.width=6, fig.height=4, dpi=300}
ggplot(aes(x=Total.Cost, y=(total.pubs)/(YearsAtISU+1)), data=subset(citations.total, Total.Cost>0 & ProfRankFac!="Other")) +
  geom_smooth(aes(ymin=pmax(0, ..ymin..)), span=.9, level=.9) +
  geom_point(shape=1, size=2, alpha=.25, guide=F) + 
  facet_wrap(~CollegeFac, scales="free", ncol=5) + 
  xlab("Startup Cost") + ylab("Yearly Publications") + 
  ggtitle("Publication Rate, By College and Years at ISU") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(trans="log10")
```

We see that in general, yearly publications are tenuously related to startup package cost. Upward trends in Engineering and Human Sciences (and the downward trend in Design) are attributable generally to one or two outliers. In LAS, there is some increase in average yearly publications initially (perhaps because of the association between Department, Startup Costs, and publication expectations), but this flattens out fairly quickly. 

### Analysis of Publications by College

```{r GScholarLAS-Arts, fig.width=6, fig.height=8, dpi=300}
ggplot(aes(x=YearsAtISU, y=jitter(total.pubs, amount=.1), shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, CollegeFac=="LAS: Arts")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding\n(Liberal Arts & Sciences: Arts)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```

```{r GScholarLAS-Science, fig.width=6, fig.height=8, dpi=300}
ggplot(aes(x=YearsAtISU, y=jitter(total.pubs, amount=.1), shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, CollegeFac=="LAS: Science")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding\n(Liberal Arts & Sciences: Sciences)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```
```{r GScholarLAS-SocialSci, fig.width=6, fig.height=8, dpi=300}
ggplot(aes(x=YearsAtISU, y=jitter(total.pubs, amount=.1), shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, CollegeFac=="LAS: Social Sci")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding\n(Liberal Arts & Sciences: Social Science)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```
In some LAS departments, like Computer Science, publication frequency does not appear to be overly impacted by career stage. In other departments, such as Chemistry, publications increase over time, and then may decrease after 5-6 years. Departments such as Music do not seem to have much publication pressure (at least not which would appear on Google Scholar). In the statistics department, most hires publish at least 2 papers per year throughout their careers, but after 3-5 years, some professors begin publishing at a much higher rate, perhaps due to increased collaboration (in fact, in the statistics department, papers with more than 6 authors are extremely uncommon until year 4-5 at ISU). 

In all departments, however, it seems that differential startup funding does not impact publication frequency or trajectory. 

```{r GScholarAGLS, fig.width=6, fig.height=7, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.pubs, shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, atISU & CollegeFac=="AGLS")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding (Ag & Life Sciences)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```

In Ag & Life Sciences, publication trends seem to be highly department-dependent. In Animal Science (and, to a lesser degree, Agronomy), publications seem to primarily increase over time. In other departments, such as NREM and ABE, publications do not seem to increase over time, but are highly variable in any given year. In all departments, it does not appear that there is a significant relationship between startup funding and publishing trajectory.


```{r GScholarBus, fig.width=6, fig.height=5, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.pubs, shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, atISU & CollegeFac=="Bus")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding (Business)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```

In the business college, publications at (on average) `r round(with(subset(citations.total, CollegeFac=="Bus"), mean(total.pubs)/mean(YearsAtISU)), 2)` publications per year. As there are relatively few hires who have been at ISU for more than 5 years, it is difficult to draw any conclusions about publishing trajectories over time. 

```{r GScholarDes, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=total.pubs, shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, atISU & CollegeFac=="Des")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding (Design)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```

The College of Design seems to have relatively few publications per person, per year (as indexed by Google Scholar), in addition to relatively few hires between 2005-2015. This scarcity of data makes drawing any concludions difficult. 

```{r GScholarEngr, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=jitter(total.pubs, amount=.1), shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, atISU & CollegeFac=="Engr")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding (Engineering)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```

The College of Engineering seems to have a constant (or perhaps slowly increasing) publication rate for individuals (with considerable variability in any single year), across all departments. The Materials Science & Engineering department is notable as there seem to be two types of hires; those who publish in great volume, and those who do not publish more than 2-3 papers per year after 5 years at ISU. 

There may be some small association between startup funding and publication trajectory within some departments, but it does not appear to be a simple relationship, and we do not have sufficient sample size within any single department to make strong conclusions.  


```{r GScholarHumSci, fig.width=6, fig.height=5, dpi=300}
ggplot(aes(x=YearsAtISU, y=jitter(total.pubs, amount=.1), shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, atISU & CollegeFac=="Hum Sci")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding (Human Sciences)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```
In the College of Human Sciences, variability in yearly publications seems to outweigh any temporal or startup funding related trends in most departments. 

```{r GScholarVet, fig.width=6, fig.height=5, dpi=300}
ggplot(aes(x=YearsAtISU, y=jitter(total.pubs, amount=.1), shape=TotalCostFac, color=TotalCostFac), 
       data=subset(citations.yearly, atISU & CollegeFac=="Vet Med")) +
  geom_line(aes(group=Name)) + 
  scale_shape_manual("Startup Funding Quantile (by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile (by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=4, override.aes=list(size=4)), color=guide_legend(ncol=4)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Yearly Publications") + 
  ggtitle("Yearly Publications and Startup Funding (Vet Med)") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_x_continuous(breaks=c(0, 5, 10), limit=c(0,10))
```

In the College of Veterinary Medicine, publications seem to be confined to certain individuals, with most individuals publishing less than 5 papers per year that were indexed by Google Scholar. It may be (particularly in VDPAM) that individuals who are awarded larger startup packages are more likely to publish many papers each year. 


