---
title: "Startup Funds Investigation"
author: "Susan VanderPlas"
date: "June 10, 2015"
output: 
  knitrBootstrap:::bootstrap_document:
    theme: cerulean
    highlight: idea
    menu: FALSE
    clean_supporting: TRUE
---


```{r setup, echo=FALSE}
library(knitr)
library(htmltools)
library(ggplot2)
# s0 <- knitr::knit_hooks$get('source')
# knitr::knit_hooks$set(
#   list(
#     source=function(x,options){
#       if(is.null(options$class)) s0(x, options)
#       else
#         tags$button(class=c())
#     }
#   )
# )
# 
knitr::opts_chunk$set(bootstrap.show.code=FALSE, bootstrap.show.output=FALSE, bootstrap.show.message=FALSE, warning=F, message=F, out.width="100%", cache=T, echo=F)
knitr::opts_knit$set(root.dir="../../IowaStateStartupFunds/")
```

# Introduction

## Goals for the project:

1. Examine trends (or lack thereof) in startup funding over time
2. Examine return on investment for startup funding 
    - Publications
    - Proposals submitted
    - Grants received
    - Tenure

## Data: 

1. Spreadsheets from the Office of the Vice President for Research
    - Name, Department, College, Salary, Start Date
    - Startup package allocation (by category) 
    - Source of startup funding
2. Database from Office of Sponsored Programs Administration (2 databases, 2005-2010 and 2010-2015. Early database will be messier)
    - Grants received (by month, as well as total award $$)
    - Proposals submitted
3. Additional information 
    - Publications (from scraping Google Scholar)
    - Where individuals came from (public/private institution, small/large)


## Other factors:

- 2003-2008 was a competitive hiring season (may have larger startup packages as a result)
- 2008 Crash and budget freeze at the state level
- American Recovery and Reinvestment Act (2009) with increased funding levels
- 2012 slowdown in research funding (may need to compare to NSF/NIH/USDA/DOD/DOE/Ames Lab funding levels)


# Analysis

## Office of the Vice President for Research Data

```{r OVPRdata}
source("Code/CleanData.R")
```

### Total Startup Package Cost
```{r TotalCost, echo=F, fig.width=9, fig.height=5, fig.cap="Total Startup Package Cost, by College"}
qplot(Year, Total.Cost, color=Dept1, data=hires, geom="jitter", size=High.Impact, shape=I(1)) +
scale_y_log10() +
scale_color_discrete(guide="none") +
scale_size_manual("High Impact Hire", values=c(2, 4)) +
ggtitle("Total Startup Package Cost")  +
theme_bw() +
facet_wrap(~College, ncol=4)
```

```{r ExpenseProportion, echo=F, fig.width=8, fig.height=5, fig.cap="Proportion of startup $ allocated for computers"}
qplot(Year, Amount/Total.Cost, color=Dept1, data=subset(startup.package, Item=="Computer.peripherals"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Proportion of package allocated to computer equipment")  +
theme_bw() +
facet_wrap(~College, ncol=4)
```

