---
title: "Startup Funds Investigation"
author: "Susan VanderPlas"
date: "June 10, 2015"
output: 
  knitrBootstrap:::bootstrap_document:
    theme: cerulean
    highlight: idea
    menu: FALSE
    clean_supporting: TRUE
---


```{r setup, echo=FALSE}
library(knitr)
library(htmltools)
library(ggplot2)
library(scales)
# s0 <- knitr::knit_hooks$get('source')
# knitr::knit_hooks$set(
#   list(
#     source=function(x,options){
#       if(is.null(options$class)) s0(x, options)
#       else
#         tags$button(class=c())
#     }
#   )
# )
# 
knitr::opts_chunk$set(bootstrap.show.code=FALSE, bootstrap.show.output=FALSE, bootstrap.show.message=FALSE, warning=F, message=F, out.width="100%", cache=T, echo=F)
knitr::opts_knit$set(root.dir="../../IowaStateStartupFunds/")
```

# Introduction

The following information was taken from meeting notes (meetings on 6/4/2015 and 5/20/2015)

## Goals for the project:

1. Examine trends (or lack thereof) in startup funding over time
2. Examine return on investment for startup funding 
    - Publications
    - Proposals submitted
    - Grants received
    - Tenure

## Data: 

1. Spreadsheets from the Office of the Vice President for Research
    - Name, Department, College, Salary, Start Date
    - Startup package allocation (by category) 
    - Source of startup funding
2. Database from Office of Sponsored Programs Administration (2 databases, 2005-2010 and 2010-2015. Early database will be messier)
    - Grants received (by month, as well as total award $$)
    - Proposals submitted
3. Additional information 
    - Publications (from scraping Google Scholar)
    - Where individuals came from (public/private institution, small/large)

## Other factors:

- 2003-2008 was a competitive hiring season (may have larger startup packages as a result)
- 2008 Crash and budget freeze at the state level
- American Recovery and Reinvestment Act (2009) with increased funding levels
- 2012 slowdown in research funding (may need to compare to NSF/NIH/USDA/DOD/DOE/Ames Lab funding levels)


# Analysis

## Office of the Vice President for Research Data

```{r OVPRdata, cache=F}
source("Code/CleanStartupFundingData.R")
```

### Total Startup Package Cost
```{r TotalCost, echo=F, fig.width=9, fig.height=5, fig.cap="Total Startup Package Cost, by College"}
qplot(Year, Total.Cost, color=Dept1, data=hires, geom="jitter", shape=I(1)) +
scale_y_log10() +
scale_color_discrete(guide="none") +
ggtitle("Total Startup Package Cost")  +
theme_bw() +
facet_wrap(~College, ncol=4)
```

### Computing Expenses
```{r ComputingExpenseProp, echo=F, fig.width=8, fig.height=5, fig.cap=c("Proportion of startup $ allocated for computers", "Amount allocated for computers")}
qplot(Year, Amount/Total.Cost, color=Dept1, data=subset(startup.package, Item=="Computer.peripherals"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Proportion of package allocated to computer equip.")  +
theme_bw() +
facet_wrap(~College, ncol=4)

qplot(Year, Amount, color=Dept1, data=subset(startup.package, Item=="Computer.peripherals"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Amount of startup funds allocated for computer equip.")  +
  scale_y_sqrt(breaks=c(0, 1000, 5000, 10000, 100000), label=c("0", "$1000", "$2500", "$10K", "$100K")) +
ylab("Amount") + 
theme_bw() +
facet_wrap(~College, ncol=4, scales="free_y")
```

In Ag & Life Science, I suspect the computer allocation has decreased due to Moore's law (it's cheaper to get computers that can process genetic data now than it was in 2005). Engineering allocations seem to be somewhat department-related (clustering of green and blue values, in particular). The business school seems remarkably standardized - the amount decreased in 2011-2013, but otherwise has been at approximately the same level (with extremely low variance). 

### Lab Space/Equipment Expenses
```{r LabExpenseProp, echo=F, fig.width=8, fig.height=5, fig.cap=c("Proportion of startup $ allocated for lab space and equipment.", "Amount allocated for lab space and equpment")}
qplot(Year, Amount/Total.Cost, color=Dept1, data=subset(startup.package, Item=="Lab.space.equipment"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Proportion of package allocated to lab space/equip.")  +
theme_bw() +
facet_wrap(~College, ncol=4)

qplot(Year, Amount, color=Dept1, data=subset(startup.package, Item=="Lab.space.equipment"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Amount of startup funds allocated for lab space/equip.")  +
theme_bw() +
facet_wrap(~College, ncol=4, scales="free_y")
```

Lab space and lab equipment are obviously only relevant for some colleges. The proportion and raw amount of funding for lab space and equipment in LAS has decreased, while the amount of funding in engineering and human sciences has increased (though the proportion of the funding package has stayed reasonably consistent).  

### Graduate Assistant Expenses
```{r GradExpenseProp, echo=F, fig.width=8, fig.height=5, fig.cap=c("Proportion of startup $ allocated for graduate assistant support.", "Amount allocated for graduate assistant support")}
qplot(Year, Amount/Total.Cost, color=Dept1, data=subset(startup.package, Item=="Graduate.assistants"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
  geom_smooth(aes(x=Year, y=Amount/Total.Cost), color=I("black"), data=subset(startup.package, Item=="Graduate.assistants")) + 
ggtitle("Proportion of package allocated for graduate assistants")  +
theme_bw() +
facet_wrap(~College, ncol=4)

qplot(Year, Amount, color=Dept1, data=subset(startup.package, Item=="Graduate.assistants"), geom="jitter", shape=I(1)) +
  geom_smooth(aes(x=Year, y=Amount), color=I("black"), data=subset(startup.package, Item=="Graduate.assistants")) + 
scale_color_discrete(guide="none") +
ggtitle("Amount of startup funds allocated for graduate assistants")  +
theme_bw() +
facet_wrap(~College, ncol=4, scales="free_y")
```

The recession seems to have hit graduate student support fairly hard (relative to other expenses). Looking at the gross amount of funding allocated for graduate students, we see that startup packages in engineering and human sciences include more support for graduate students than in the past (this is particularly dramatic in human sciences, though the amount of funding is still overall lower than in engineering). In the other colleges, the amount of support for graduate students has stayed relatively consistent over the past 10 years. 

### Summer Support
```{r SummerExpenseProp, echo=F, fig.width=8, fig.height=5, fig.cap="Proportion of startup $ allocated for summer support."}
qplot(Year, Amount/Total.Cost, color=Dept1, data=subset(startup.package, Item=="Summer.support"), geom="jitter", shape=I(1)) +
  geom_smooth(aes(x=Year, y=Amount/Total.Cost), color=I("black"), data=subset(startup.package, Item=="Summer.support")) + 
scale_color_discrete(guide="none") +
ggtitle("Proportion of package allocated for summer support")  +
theme_bw() +
facet_wrap(~College, ncol=4)

qplot(Year, Amount, color=Dept1, data=subset(startup.package, Item=="Summer.support"), geom="jitter", shape=I(1)) +
  geom_smooth(aes(x=Year, y=Amount), color=I("black"), data=subset(startup.package, Item=="Summer.support")) + 
scale_color_discrete(guide="none") +
ggtitle("Amount of startup funds allocated for summer support")  +
theme_bw() +
facet_wrap(~College, ncol=4, scales="free_y")
```

Summer funding seems to be fairly common in some areas (business, some of LAS) and almost unheard of in others (design, vet med). 

### Research Support
```{r ResearchExpenseProp, echo=F, fig.width=8, fig.height=5, fig.cap=c("Proportion of startup $ allocated for research support.", "Amount of startup $ allocated for research support.")}
qplot(Year, Amount/Total.Cost, color=Dept1, data=subset(startup.package, Item=="Research.support"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Proportion of package allocated for research support")  +
theme_bw() +
facet_wrap(~College, ncol=4)

qplot(Year, Amount, color=Dept1, data=subset(startup.package, Item=="Research.support"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Amount of funding allocated for research support")  +
theme_bw() +
facet_wrap(~College, ncol=4, scales="free_y")
```

### Moving Expenses
```{r MovingExpenseProp, echo=F, fig.width=8, fig.height=5, fig.cap=c("Proportion of startup $ allocated for moving expenses.", "Amount allocated to moving expenses")}
qplot(Year, Amount/Total.Cost, color=Dept1, data=subset(startup.package, Item=="Moving.expenses"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Proportion of package allocated for moving expenses")  +
theme_bw() +
facet_wrap(~College, ncol=4)

qplot(Year, Amount, color=Dept1, data=subset(startup.package, Item=="Moving.expenses"), geom="jitter", shape=I(1)) +
scale_color_discrete(guide="none") +
ggtitle("Amount allocated for moving expenses")  +
theme_bw() +
facet_wrap(~College, ncol=4)
```

Moving expenses amounts seem to be reasonably consistent across colleges. The business college seems to once again have a standard package that may change year to year with occasional negotiation. The same may be true of the vet med college. 

## Office of Sponsored Programs Administration Data

Datasets from OSPA contain information on proposals and awards from 2005-2010 and 2010-2015 (4 total databases). 


```{r OSPAdata, cache=F}
source("Code/CleanAwardsData.R")
source("Code/CleanProposalsData.R")
```

### Proposals

```{r ProposalExploration}

prop$Source2 <- factor(prop$Source, labels=c("FY06-FY10", "FY11-FY16"), levels=c("Proposals FY06 to FY10.csv", "KC Investigator Proposal ALL 06072015.csv"))

qplot(Activity.Type, data=prop, color=I("black"), fill=I("grey")) + facet_grid(Source2~.) + theme_bw()  + theme(legend.position="bottom", axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Proposal Activities")

qplot(Proposal.Type, data=prop, color=I("black"), fill=I("grey")) + facet_grid(Source2~.) + theme_bw()  + theme(legend.position="bottom", axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Proposal Type")

qplot(Grant.Type, data=prop, color=I("black"), fill=I("grey")) + facet_grid(Source2~.) + theme_bw()  + theme(legend.position="bottom", axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Funding Type")

qplot(Proposal.Status, fill=Source, position="stack", data=prop) + theme_bw() + 
  theme(legend.position="bottom")  +
  ggtitle("Proposal Status")

prop.unique <- prop %>% group_by(Proposal.Number, Submit.Date, Source2) %>% summarise(ct=length(Proposal.Number))

qplot(floor_date(Submit.Date, "month"), color=I("black"), fill=Source2, position="stack", data=prop.unique) + theme_bw() + theme(legend.position="bottom") + xlab("Submission Date") + ylab("Count") + ggtitle("Number of Proposals Submitted")

rm(prop.unique)

prop.span <- melt(prop, id.vars = c("Proposal.Number", "Source2", "Proposal.Title"), measure.vars=c("Start.Date", "End.Date", "Submit.Date"), value.name = "Date", variable.name = "Proposal.Span")
prop.span$Proposal.Span <- factor(prop.span$Proposal.Span, levels=c("Submit.Date", "Start.Date", "End.Date"))

ggplot(data=prop.span) + 
  geom_histogram(aes(x=floor_date(Date, "quarter"), fill=Source2), 
               position="stack", adjust=1.25, trim=T, binwidth=15*60*24*365) + 
  facet_grid(Proposal.Span~.) + 
  theme_bw() + 
  scale_x_datetime(limits=c(mdy("07012001"), mdy("07012026")), breaks=date_breaks("5 years")) + 
  scale_fill_discrete("Data Source") + xlab("Date") + ylab("Density") + ggtitle("Proposal Timeline") + 
  theme(legend.position="bottom") 

prop.span <- melt(prop, id.vars = c("Proposal.Number", "Source2", "Proposal.Title", "Submit.Date"), measure.vars=c("Start.Date", "End.Date"), value.name = "Date", variable.name = "Proposal.Span")
prop.span$Proposal.Span <- factor(prop.span$Proposal.Span, levels=c("Start.Date", "End.Date"))
prop.span$FundingLag <- with(prop.span, difftime(Date, Submit.Date, units="days")) 
prop.span$FundingLag <- as.numeric(prop.span$FundingLag)/(365.25) # convert to years

ggplot(data=prop.span) + 
  geom_histogram(aes(x=FundingLag), fill="grey", color="black", binwidth=.125, inherit.aes=F) + 
  facet_wrap(~Proposal.Span, scales="free_x") + 
  theme_bw() + 
  scale_x_continuous(limits=c(-1, 10), breaks=0:9) + 
  xlab("Years from Proposal Submission Date") + ylab("Count") + ggtitle("Proposal Timeline") + 
  theme(legend.position="bottom") 

rm(prop.span)

ggplot(data=prop) + 
  geom_histogram(aes(x=as.numeric(difftime(End.Date, Start.Date, units="days")/365.25)), 
                 binwidth=.25, color="black", fill="grey") + 
  scale_x_continuous("Proposal Duration", limits=c(0, 10), breaks=0:10) + 
  ylab("Count") + ggtitle("Duration of Proposed Funding") + theme_bw() + 
  theme(legend.position="bottom") 

```

### Awards

```{r AwardExploration}

awards$Source2 <- factor(awards$Source, labels=c("FY06-FY10", "FY11-FY16"), levels=c("Awards FY06 to FY10.csv", "KC Investigator Award ALL 06072015.csv"))

qplot(Activity.Type, data=awards, color=I("black"), fill=I("grey")) + facet_grid(Source2~.) + theme_bw()  + theme(legend.position="bottom", axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Funding Activity Types")

qplot(Award.Type, data=awards, color=I("black"), fill=I("grey")) + facet_grid(Source2~.) + theme_bw()  + theme(legend.position="bottom", axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Funding Award Type")

award.span <- melt(awards, id.vars = c("Award.Number", "Source2", "Award.Title"), measure.vars=c("Start.Date", "End.Date"), value.name = "Date", variable.name = "Award.Span")
award.span$Award.Span <- factor(award.span$Award.Span, levels=c("Start.Date", "End.Date"))

ggplot(data=award.span) + 
  geom_histogram(aes(x=floor_date(Date, "quarter"), fill=Source2), 
               position="stack", adjust=1.25, trim=T, binwidth=15*60*24*365) + 
  facet_grid(Award.Span~.) + 
  theme_bw() + 
  scale_x_datetime(limits=c(mdy("07012001"), mdy("07012026")), breaks=date_breaks("5 years")) + 
  scale_fill_discrete("Data Source") + xlab("Date") + ylab("Count") + ggtitle("Award Timeline")

rm(award.span)

ggplot(data=awards) + 
  geom_histogram(aes(x=as.numeric(difftime(End.Date, Start.Date, units="days"))/365.25), 
                 color="black", fill="grey", binwidth=.25) + 
  theme_bw() + 
  scale_x_continuous("Years of Funding", limits=c(-1, 10), breaks=0:9) + 
  scale_fill_discrete("Data Source") + xlab("Date") + ylab("Count") + ggtitle("Funding Span") + theme(legend.position="bottom")

qplot(Award.Status, data=awards, color=I("black"), fill=I("grey")) + facet_wrap(~Source2) + theme_bw() + theme(legend.position="bottom", axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Award Status")

awards$novalue <- c("No Monetary Value", "Monetary Value")[1+as.numeric(awards$Award.Amount>0)]
qplot(Award.Status, fill=novalue, position="stack", data=awards, facets=~Source2) + theme_bw() + theme(legend.position="bottom", axis.text.x=element_text(angle=45, hjust=1)) + ggtitle("Award Status by Grant Amount")


ggplot(data=awards) + 
  geom_histogram(aes(x=Start.Date, weight=Award.Amount), 
                 color="black", fill="grey") + 
  theme_bw() + 
  scale_fill_discrete("Data Source") + xlab("Date") + ylab("Count") + ggtitle("Funding Span") + theme(legend.position="bottom")
```




# Merged Data

```{r MergeDatasets, cache=F}
received.money <- awards %>% 
  group_by(Investigator) %>% 
  summarize(grant.received = sum(Award.Amount), 
            n.awards = length(unique(Award.Number)))
received.money$grant.received[is.na(received.money$grant.received)] <- 0

submitted.proposals <- prop %>% 
  group_by(Investigator) %>% 
  summarize(total.proposal.money = sum(Proposal.Amount), 
            n.proposals = length(unique(Proposal.Number)), 
            prop.fund.pct = mean(Proposal.Status=="Funded"))
submitted.proposals$total.proposal.money[is.na(submitted.proposals$total.proposal.money)] <- 0

tmp <- merge(hires, received.money, by.x="Name", by.y="Investigator", all.x=T, all.y=F)
tmp <- merge(tmp, submitted.proposals, by.x="Name", by.y="Investigator", all.x=T, all.y=F)
tmp$grant.received[is.na(tmp$grant.received)] <- 0
tmp$total.proposal.money[is.na(tmp$total.proposal.money)] <- 0

tmp$ProfRankFac <- tmp$ProfRank %>% 
  str_replace("(Adjunct/Lecturer)|(Affiliate Prof)|(Clinician)", "Other") %>%
  str_replace("Distg ", "") %>%
  factor(levels=c("Asst Prof", "Assoc Prof", "Prof", "Other"), labels=c("Assistant", "Associate", "Professor", "Other"))

tmp$CollegeFac <- tmp$College %>% factor(labels=c("AGLS", "Bus", "Des", "Engr", "Hum Sci", "LAS", "Vet Med"))

tmp$YearsAtISU <- 2015 - tmp$Year

# Transformation - log(x+1) to deal with zero data 
require(scales)
log_one_trans <- function() trans_new("log_one", function(x) log10(x+1), function(x) 10^x-1, domain=c(0, Inf))

```

We'll begin by looking at the timeline - how many years at ISU are necessary before grant applications are successful? We split this by college and faculty rank (as we'd expect that full professors who are hired should be able to command grant money sooner than new faculty). Subsequent graphs show data from hired professors (i.e. not adjuncts, affiliates, lecturers, or clinicians). 

```{r, fig.width=6, fig.height=8, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, color=CollegeFac, shape=CollegeFac), data=subset(tmp, ProfRankFac!="Other" )) +
  geom_jitter(size=4, alpha=.5) + 
  # geom_smooth(se=F, adjust=2) + 
  scale_shape_manual(values=c('A', 'B', 'D', 'E', 'H', 'L', 'V'), guide=F) +
  scale_color_discrete(guide=F) + 
  # guides(shape=guide_legend(ncol=7), color=guide_legend(ncol=7)) + 
  facet_grid(CollegeFac~ProfRankFac, scales="free") + 
  xlab("Years from Hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 1e3, 1e5, 1e7))
```
The most noticeable difference between colleges is that in engineering, assistant professors get grants about five years after they are hired, and it is very rare for professors at that stage in their careers not to get grant funding. This trend is present (though less pronounced) in Ag & Life Sciences and Human Sciences as well, though some assistant professors in those colleges seem to get funding much earlier in their careers (perhaps because of large grants given to groups of professors). In Liberal Arts and Sciences, grant funding does seem to increase with experience for some hires, but perhaps because LAS includes both arts (which are less grant-reliant) and sciences, this trend is less pronounced than in other colleges. 

In order to examine the effect of startup funding on grant receipts over a career of variable duration, we will examine the distribution of startup funding by college, categorizing startup costs by the quartile (calculated for the hire's college). Below, vertical lines mark these quantiles

```{r, fig.width=6, fig.height=6, dpi=300}
tmp$YearsAtISUFac <- cut_number(tmp$YearsAtISU, n=4)

startup.quantiles <- tmp %>% group_by(CollegeFac) %>% summarize(Q1=quantile(Total.Cost, .25, type=8), Q2=quantile(Total.Cost, .5, type=8), Q3=quantile(Total.Cost, .75, type=8), Q4=quantile(Total.Cost, 1, type=8)) %>% melt(id.vars=1, measure.vars=2:ncol(.), variable.name="Quantile", value.name="Total.Cost")

tmp <- tmp %>% group_by(CollegeFac) %>% mutate(TotalCostFac = as.numeric(cut(Total.Cost, breaks=quantile(Total.Cost, probs=0:4/4))))
tmp$TotalCostFac[is.na(tmp$TotalCostFac)] <- 1
tmp$TotalCostFac <- factor(tmp$TotalCostFac, levels=1:4, labels=c("Startup $: Q1", "Startup $: Q2", "Startup $: Q3", "Startup $: Q4"))

ggplot(aes(x=Total.Cost, fill=CollegeFac), data=tmp) + 
  geom_density(aes(y=..density..), origin=0, color="black", trim=T) +
  geom_vline(aes(xintercept=Total.Cost), data=startup.quantiles) + 
  scale_fill_discrete(guide=F) + 
  scale_x_continuous(trans='sqrt', breaks=c(1000, 25000, 50000, 100000, 250000, 500000, 1e6, 2e6), labels=c("1K", "25K", "50K", "100K", "250K", "500K", "1M", "2M"), limits=c(0, 2e6)) + 
  facet_grid(CollegeFac~.) + 
  ggtitle("Distribution of Startup Costs by College") +
  xlab("Startup Package Cost") + ylab("Count") + 
  # guides(fill=guide_legend(ncol=7)) + 
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical")
```

Using these quantile calculations, we then can examine total grants received as explained by years since hired. 

```{r, fig.width=8, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(tmp, ProfRankFac!="Other" )) +
  geom_jitter(size=3) + 
  geom_smooth(method="lm", span=.5, se=F, fullrange=F, aes(x=YearsAtISU, y=grant.received)) + 
  scale_shape_manual("Startup Funding Quantile\n(by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile\n(by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=7, override.aes=list(size=4)), color=guide_legend(ncol=7)) + 
  facet_wrap(~CollegeFac, ncol=4) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) +
  scale_x_continuous(breaks=c(0, 5, 10))
```

It is important to note several caveats at this point: first, hires are not separated out by faculty rank or position, so hires whose primary responsibilities include administrative work may command large startup packages but bring in relatively little grant funding. Additionally, different colleges have different funding structures: LAS, for instance, may need to provide lab startup funding for science based departments, but would not have to provide this funding for new hires in English or History. Finally, the x-axis shows years since hire, which does not necessarily translate to years at ISU (faculty hired in 2005 may have left in 2011 after receiving tenure, for example). 

Linear regression lines are provided here as (extremely) rough approximations; it may be useful to examine the right-most endpoint (rather than the entire line) as a prediction of total grant money received after 10 years at ISU. 

With these caveats in mind, 

* Startup funding in LAS is correlated with later grant funding; this is likely explained by departmental startup funding (and grant funding) norms. 
* Startup funding in engineering is largely not predictive of grant receipts; after 10 years, hires with small startup packages receive approximately the same grant funding as hires with large startup packages. 
* Startup funding in Human Sciences is correlated with later grant funding- this may again be a departmental norm situation (psychologists may require lab space and compete for NSF/NIH grants, where professors in the education department require fewer financial resources when hired and receive fewer grants.)
* Grants are relatively rare in the business college.   

Examining the LAS results in more detail, we see that departmental startup packages largely fall in the same quantiles. In Chemistry, Physics, BBMB, EEOB, Computer Science, Statistics, Economics, and Math, the career trajectory of grant funding suggests that the large startup packages are worthwhile investments; these disciplines are also largely in the top two quartiles for startup funding. Psychology is more of a mixed bag; many psychologists received large startup packages but have not received significant grant funding. Disciplines with more economical startup packages are also less likely to have faculty receiving large grants; this is likely due to discipline norms, rather than faculty quality. 
```{r, fig.width=6, fig.height=6, dpi=300}
ggplot(aes(x=YearsAtISU, y=grant.received, shape=TotalCostFac, color=TotalCostFac), data=subset(tmp, CollegeFac=="LAS" & ProfRankFac!="Other" )) +
  geom_jitter(size=3) + 
  scale_shape_manual("Startup Funding Quantile\n(by college)", values=as.character(1:4), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  scale_color_manual("Startup Funding Quantile\n(by college)", values=c("purple", "blue", "darkgreen", "orange"), labels=c("Q1", "Q2", "Q3", "Q4")) + 
  guides(shape=guide_legend(ncol=7, override.aes=list(size=4)), color=guide_legend(ncol=7)) + 
  facet_wrap(~Dept1) + 
  xlab("Years since hire") + ylab("Total Grant $ Received") + 
  ggtitle("Grant receipts based on hire date") +
  theme_bw() + theme(legend.position="bottom", legend.direction="vertical") + 
  scale_y_continuous(trans="log_one", breaks=c(0, 10, 1e3, 1e5, 1e7), limits=c(0, 2e7)) + 
  scale_x_continuous(breaks=c(0, 5, 10))
```

